Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-05-24 doc/bcspaper/article.tex
Maria edits all sections

\usepackage{subcaption}

\captionsetup{compatibility=false}

@@

\title*{Partridge: An effective system for the automatic classification of the types of academic papers}
\titlerunning{Partridge: classifying academic papers}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{James Ravenscroft \and Maria Liakata \and Amanda Clare}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{James Ravenscroft, Amanda Clare \at Department of Computer Science, Aberystwyth University, Aberystwyth, SY23 3DB, UK  \email{jrr9@aber.ac.uk}
\and Maria Liakata \at Name, Address of Institute \email{name@email.address}}

@@

\subsection*{Collection of scientific articles} Partridge allows users to
upload any paper which is free of copyright restrictions.  For the purpose of
this study we needed a large set of papers that we could label with CoreSC to
investigate how this information assists classification into paper type.  Open
Access (OA) journals provide free read access to their papers, but many do not
permit the user of articles for data mining purposes. The PLoS
journals\footnote{PLoS: \url{http://www.plos.org/}} contain large volumes of OA
literature under a permissive license that allows data mining.  They also use
the Pubmed Central markup schema, which is compatible with SAPIENTA, for papers
published through their journal. The PLoS journals advanced search offers
approximately 50 types of paper through which to restrict the search.  Many of
these paper type categories contain too few papers to be useful, others are too
specific (e.g.``Message from PLoS'') or appear rather adhoc. Indeed it is not
clear how these paper types have been identified. We chose to look at a range
of types, some of which we expect to overlap or have an unclear distinction.
These are namely: Essay, Correspondence, Case Study, Synopsis, Perspective,
Viewpoint, Opinion, Review, Research.

% in conclusions add how this can give recommendations for Plos and other journals 
The papers were downloaded from PLoS using {\em plosget.py}, a short python
script that uses the PLoS RESTful search API. The number of papers downloaded
per paper type category was as follows: 200 Essay, 99 Correspondence, 174
Synopsis, 200 Perspective, 74 Viewpoint, 93 Opinion, 107 Case Study, 312
Review, 200 Research.  These formed a corpus of 1459 papers.

@@

\begin{table}
\caption{Per-class recall, precision and F-measure averaged over 10-fold cross validation, reporting the results on the held-out validation segment}
\label{tab:recallPrecision}       % Give a unique label
\begin{tabular}{p{2cm}p{2.4cm}p{2cm}p{4.9cm}}
\hline\noalign{\smallskip}
Classes & Recall & Precision & F-measure  \\
\noalign{\smallskip}\svhline\noalign{\smallskip}
Case Study   &     0.243    &    0.245     &   0.244 \\
Correspondence  &      0.545    &    0.505    &    0.524 \\ 
Essay     &   0.450     &   0.462    &    0.456 \\
Opinion   &     0.247    &    0.235  &      0.241 \\   
Perspective &       0.255  &      0.304  &       0.277 \\
Research     &   0.700    &    0.619    &    0.657 \\
Review      &  0.631     &   0.625     &   0.628 \\
Viewpoint   &     0.149     &   0.157  &      0.153 \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
\end{table}

A confusion matrix for the paper types in given in Table \ref{tab:confusion}. Research papers are usually predicted to belong to the Research class, and are sometimes confused with Case Study. However, Case Study papers are often predicted to be Research or Review. There are fewer Case Study papers for the learning algorithm to train on, and many research papers, so we would expect this outcome. We might also expect that Research and Case Study would share similar paper structures. 
We initially expected the classes Essay, Opinion, Perspective and Viewpoint to be confused, as the four labels all indicate a paper containing an author's personal thoughts on an issue, rather than traditional experiment-driven science. Opinion is confused with Perspective and Viewpoint. Perspective is confused with almost all classes, except Research. Viewpoint is a small class and mostly confused with Opinion and Perspective. From this it would seem that there is a clear grouping of paper types which are very similar, and could better be labelled as a single type: \{Opinion, Perspective, Viewpoint\}. However, Essay appears to be different.

Essay is confused with Perspective and Review. Review is confused with Case Study and Essay, and a few others. They share overlapping aspects, but are not entirely synonymous. A review should cover a wide range of background material (which we might also expect to find in an Essay), and might make recommendations based on the material presented (which we might also expect to find in a Case Study).

@@

To summarise, we have demonstrated that paper type can largely be predicted from the distribution of the CoreSC sentence labels in the full text of a paper. The CoreSC labels themselves are automatically annotated. We have described some of the particular CoreSC features that determine a paper type, and discussed which of the paper types are not easily separable in this way. Analysis of an example tree shows the decision making process to be complex, but agrees with our general understanding of paper types. The Partridge online system now allows refinement of paper search using CoreSC and paper type, both of which can be intelligently determined using machine learning methods.

@@

\begin{acknowledgement}
Thanks to the Leverhulme for Maria's funding....
\end{acknowledgement}

>>>>

%\usepackage{subcaption}

%\captionsetup{compatibility=false}

@@

\title*{Partridge: An effective system for the automatic classification of the types of academic papers}
\titlerunning{Partridge: classifying academic papers}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{James Ravenscroft \and Maria Liakata \and Amanda Clare}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{James Ravenscroft, Amanda Clare \at Department of Computer Science, Aberystwyth University, Aberystwyth, SY23 3DB, UK  \email{jrr9@aber.ac.uk}
\and Maria Liakata \at Department of Computer Science, University of Warwick \email{M.Liakata@warwick.ac.uk}}

@@

\subsection*{Collection of scientific articles} Partridge allows users to
upload any paper which is free of copyright restrictions.  For the purpose of
this study we needed a large set of papers that we could label with CoreSC to
investigate how this information assists classification into paper type.  Open
Access (OA) journals provide free read access to their papers, but many do not
permit the user of articles for data mining purposes. The PLoS
journals\footnote{PLoS: \url{http://www.plos.org/}} contain large volumes of OA
literature under a permissive license that allows data mining.  They also use
the Pubmed Central markup schema, which is compatible with SAPIENTA, for papers
published through their journal. The PLoS journals advanced search offers
approximately 50 types of paper through which to restrict the search.  Many of
these paper type categories contain too few papers to be useful, others are too
specific (e.g.``Message from PLoS'') or appear rather adhoc. Indeed it is not
clear how these paper types have been identified. We chose to look at a range
of types, some of which we expect to overlap or have an unclear distinction.
These are namely: Essay, Correspondence, Case Study, Perspective,
Viewpoint, Opinion, Review, Research.

% in conclusions add how this can give recommendations for Plos and other journals 
The papers were downloaded from PLoS using {\em plosget.py}, a short
python script that uses the PLoS RESTful search API. The number of
papers downloaded per paper type category was as follows: 200 Essay,
99 Correspondence, 107 Case Study, 200 Perspective, 74 Viewpoint, 93
Opinion, 312 Review, 200 Research.  These formed a corpus of 1285 %1459-174 after removing synopsis. Should double check the numbers add up
papers.

@@

\begin{table}
\caption{Per-class recall, precision and F-measure micro-averaged over 10-fold cross validation, reporting the results on the held-out validation segment}
\label{tab:recallPrecision}       % Give a unique label
\begin{tabular}{p{2cm}p{2.4cm}p{2cm}p{4.9cm}}
\hline\noalign{\smallskip}
Classes & Recall & Precision & F-measure  \\
\noalign{\smallskip}\svhline\noalign{\smallskip}
Case Study   &     0.243    &    0.245     &   0.244 \\
Correspondence  &     {\bf  0.545}    &    {\bf 0.505}    &    {\bf 0.524} \\ 
Essay     &   0.450     &   0.462    &    0.456 \\
Opinion   &     0.247    &    0.235  &      0.241 \\   
Perspective &       0.255  &      0.304  &       0.277 \\
Research     &   {\bf 0.700}    &    {\bf 0.619}    &    {\bf 0.657} \\
Review      &  {\bf 0.631}     &   {\bf 0.625}     &   {\bf 0.628} \\
Viewpoint   &     0.149     &   0.157  &      0.153 \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
\end{table}

A confusion matrix for the paper types in given in Table
\ref{tab:confusion}. Research papers are usually predicted to belong
to the Research class, and are sometimes confused with Case
Study. However, Case Study papers are often predicted to be Research
or Review. The ratio between Case study, Research and Review papers is
~1:2:3 so the outcome is not entirely surprising in terms of the data
size effect. We expected Research and Case Study to share similar
paper structures but perhaps the fact that Case Study is equally
confused with Research and Review suggests two distinct types of Case
Study.  We expected the classes Essay, Opinion, Perspective and
Viewpoint to be confused, as the four labels all indicate a paper
containing an author's personal thoughts on an issue, rather than
experiment-driven science. Opinion is confused with Perspective and
Viewpoint. Perspective is confused with almost all classes, except
Research. Viewpoint is a small class and mostly confused with Opinion
and Perspective. From this it would seem that the paper types
\{Opinion, Perspective, Viewpoint\} are very similar and should be
grouped together into one category. Indeed when we trained the model
on a super class consisting of the previous three categories,
performance was improved overall with the following f-measures:
OpinionSuper: 59, Research: 67.6, Review: 63.7, Essay: 44.9,
Correspondence: 46 abd Case Study 24.8.

Essay appears to be distinct enought to justify a class of its own.
Essay is confused with Perspective and Review. Review is the second
best category but is at times confused with Case Study and Essay. A
possible explanation is that Review should cover a wide range of
background material (which we may also expect to find in an Essay),
and could make recommendations based on the material presented (which
we might also expect to find in a Case Study).

@@

To summarise, we have demonstrated that paper type can largely be
predicted from the distribution of the automatically obtained CoreSC
sentence labels in the full text of a paper. We have described some of
the particular CoreSC features that determine a paper type
(e.g. Corresponsence characterised by very little Background, Opinions
characterised by low percentage of Experiment) and discussed which of
the paper types are not easily separable in this way. We recommended
for example merging the Opinion, Viewpoint and Perspective articles
into a single category, which increased overall
classification performance. It is also potentially useful to distinguish between two types of Case study.
Analysis of an example tree shows the decision making
process to be complex, but agrees with our general understanding of
paper types. The Partridge online system now allows refinement of
paper search using CoreSC and paper type, both of which can be
intelligently determined using machine learning methods.

@@

\begin{acknowledgement}
Dr Liakata is in receipt of a Leverhulme trust Early Career Fellowship.
\end{acknowledgement}
